---
title: "Custom Scales"
output: github_document
---

```{r include = FALSE}
library(ggplot2)
library(dplyr)
data(airquality)
tall_table = data.frame("Day" = c(1, 1, 2, 2, 3, 3), "Variable" = c("Ozone", "Temp", "Ozone", "Temp", "Ozone", "Temp"), "Value" = c(41, 67, 36, 72, 12, 74))
```

Typically ggplot likes "tall" data. That is, if you want to plot 2 time series on the same plot, you're supposed to have a column that denotes what time series is being plotted, and one that denotes the value to plot, like this:
```{r echo = FALSE}
tall_table
```

Then when plotting a time series you can do something simple like
```ggplot(data) + geom_line(aes(x = Day, y = Value, colour = Variable))```

but this is not always the format your data will come in, or practical. Often you'll have "wide" data, where each value has it's own column
```{r echo = FALSE}
head(airquality)
```
You can plot multiple lines on top of each other, by simply adding more `geom_line()`s but this will not automatically make a legend for your graph.
To do that you should set up a manual scale, like this.

```{r}
ggplot(airquality %>% filter(Month == 5)) +
  geom_line(mapping = aes(x = Day, y = Wind, colour = "1"), linewidth = 1) +
  geom_line(mapping = aes(x = Day, y = Temp, colour = "2"), linewidth = 1) +
  labs(x = "Day", y = "Value") +
  scale_colour_manual(name = "Air Variable", values = c("#E16A86", "#00AD9A"), labels = c("Wind", "Temperature"))
```

Don't forget to set the labels too, otherwise the y axis will probably have the wrong name.
This also applies to any other type of graph where you might want to layer geoms on top of each other.

## A second scale

If you want to add a second scale, for example if you want to also have a different line type, then you can simple add a new `scale_x_manual` geom. If the new scale has the same name and the same labels, it will overlay into a single legend.
```{r}
ggplot(airquality %>% filter(Month == 5)) +
  geom_line(mapping = aes(x = Day, y = Wind, colour = "1", linetype = "1"), linewidth = 1) +
  geom_line(mapping = aes(x = Day, y = Temp, colour = "2", linetype = "2"), linewidth = 1) +
  labs(x = "Day", y = "Value") +
  scale_colour_manual(name = "Air Variable", values = c("#E16A86", "#00AD9A"), labels = c("Wind", "Temperature")) +
  scale_linetype_manual(name = "Air Variable", values = c("solid", "dotted"), labels = c("Wind", "Temperature"))
```